The dynamics of starvation and recovery

### 2017-11-03 NatureComm_REVISION/YKR_NatureComm-Revision-1.tex
revision 3 figure changes

% \noindent {\bf Role of allometry} \\
%[Link Allometry stuff to our model - what does it provide to the story]
%[Introduce biological and linked constraints]
%[Allometries capture vast amounts of diversity via a single parameter --- body size]
%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
\noindent {\bf Results \& Discussion}
\noindent \paragraph*{{\bf Allometry minimizes extinction risk.}} While there are no {\it a priori} constraints on the parameters in the NSM, we expect that each species should be restricted to a distinct portion of the parameter space.  We
use allometric scaling relations to constrain the covariation of rates in a
principled and biologically meaningful manner (see Methods).  Allometric scaling relations
highlight common constraints and average trends across large ranges in body
size and species diversity. Many of these relations can be derived from a
small set of assumptions and in the Methods we describe our framework to determine the
covariation of timescales and rates across a range of body sizes for each of
the key parameters of our model (cf. ref.~\citep{Yodzis:1992hg}).  
% We are thereby able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.

@@

% \noindent {\bf Energetic barriers to body size}
\noindent \paragraph*{{\bf Predicting energetic equivalence and body size limits.}} The NSM correctly predicts that species with smaller masses have larger
steady-state population densities (Fig.~\ref{fig:mass}a).  Similar
predictions have been made for carnivore populations using alternative
consumer-resource models \citep{DeLong:2012kw}.  Moreover, we show that the
NSM provides independent theoretical support for the energy equivalence
hypothesis and Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998}.  The
energy equivalence hypothesis argues that the total energy use,
$B_{\rm tot}$, of a population is constant independent of species
size~\citep{Damuth:1987kr,allen2002,enquist1998}. %Previous studies have focused on t
This hypothesis is based on observations showing that the steady state
abundance, $N^*$, of a species is proportional to the inverse of individual
metabolism, such that
$N^*\propto M^{-3/4}/B_{0}$~\citep{allen2002,enquist1998}.  This relationship
implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been
shown to hold in both mammalian and vascular plant communities
\citep{Damuth:1987kr,allen2002,enquist1998}.  Figure \ref{fig:mass}a shows
that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of
organism sizes and Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly
constant over this same range.  This result is remarkable because it
illustrates that the steady state values of the NSM combined with the derived
timescales naturally give rise to energy equivalence.

@@

\noindent {\bf Conclusion} \\
{\color{red} Complementary previous work has analyzed the tradeoffs associated with carnivory in terms of the energy expenditure on hunting and prey size, where a maximum carnivore size between 700 and 1,000 is predicted from considerations of energy intake going to zero \cite{Carbone:1999ju}. Similarly, it has been shown in the past that the largest body size within a particular lineage should be a function of the metabolic normalization constant and a critical death rate as supported by data. This approach, which consider the energetics associated with active metabolism, provides a complementary prediction for maximum size from energetics \cite{Okie:2013ju}, and also suggests that future connections between the scaling of fat and muscle mass should be systematically connected with $B_{0}$.
}

>>>>

% \noindent {\bf Role of allometry} \\
%[Link Allometry stuff to our model - what does it provide to the story]
%[Introduce biological and linked constraints]
%[Allometries capture vast amounts of diversity via a single parameter --- body size]
%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
\noindent {\bf Results}
\noindent \paragraph*{{\bf The allometry of extinction risk.}} While there are no {\it a priori} constraints on the parameters in the NSM, we expect that each species should be restricted to a distinct portion of the parameter space.  We
use allometric scaling relations to constrain the covariation of rates in a
principled and biologically meaningful manner (see Methods).  Allometric scaling relations
highlight common constraints and average trends across large ranges in body
size and species diversity. Many of these relations can be derived from a
small set of assumptions and in the Methods we describe our framework to determine the
covariation of timescales and rates across a range of body sizes for each of
the key parameters of our model (cf. ref.~\citep{Yodzis:1992hg}).  
% We are thereby able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.

@@

% \noindent {\bf Energetic barriers to body size}
\noindent \paragraph*{{\bf Predicting Damuth's Law and body size limits.}} The NSM correctly predicts that species with smaller masses have larger
steady-state population densities (Fig.~\ref{fig:mass}a).  Similar
predictions have been made for carnivore populations using alternative
consumer-resource models \citep{DeLong:2012kw}.  Moreover, we show that the
NSM provides independent theoretical support for the energy equivalence
hypothesis and Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998}.  The
energy equivalence hypothesis argues that the total energy use,
$B_{\rm tot}$, of a population is constant independent of species
size~\citep{Damuth:1987kr,allen2002,enquist1998}. %Previous studies have focused on t
This hypothesis is based on observations showing that the steady state
abundance, $N^*$, of a species is proportional to the inverse of individual
metabolism, such that
$N^*\propto M^{-3/4}/B_{0}$~\citep{allen2002,enquist1998}.  This relationship
implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been
shown to hold in both mammalian and vascular plant communities
\citep{Damuth:1987kr,allen2002,enquist1998}.  Figure \ref{fig:mass}a shows
that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of
organism sizes and Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly
constant over this same range.  This result is remarkable because it
illustrates that the steady state values of the NSM combined with the derived
timescales naturally give rise to energy equivalence.

@@

\noindent {\bf Discussion} \\
{\color{red} Complementary previous work has analyzed the tradeoffs associated with carnivory in terms of the energy expenditure on hunting and prey size, where a maximum carnivore size between 700 and 1,000 is predicted from considerations of energy intake going to zero \cite{Carbone:1999ju}. Similarly, it has been shown in the past that the largest body size within a particular lineage should be a function of the metabolic normalization constant and a critical death rate as supported by data. This approach, which consider the energetics associated with active metabolism, provides a complementary prediction for maximum size from energetics \cite{Okie:2013ju}, and also suggests that future connections between the scaling of fat and muscle mass should be systematically connected with $B_{0}$.
}

### 2017-11-03 NatureComm_REVISION/YKR_NatureComm-Revision-2.tex
revision 3 figure changes

Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
Many mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development until nutritional reserves can be accumulated~\citep{Mead:1989dt,Sandell:1990kw}.
Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\citep{Bulik:1999eo,Trites:2003cc}.
In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \citep{Glazier:2009hq}.
The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the universality of the fundamental tradeoff between somatic and reproductive investment.

@@

% \noindent {\bf Energetic barriers to body size}
\noindent \paragraph*{{\bf Predicting Damuth's Law and body size limits.}} The NSM correctly predicts that species with smaller masses have larger
steady-state population densities (Fig.~\ref{fig:mass}a).  Similar
predictions have been made for carnivore populations using alternative
consumer-resource models \citep{DeLong:2012kw}.  Moreover, we show that the
NSM provides independent theoretical support for the energy equivalence
hypothesis and Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998}.  The
energy equivalence hypothesis argues that the total energy use,
$B_{\rm tot}$, of a population is constant independent of species
size~\citep{Damuth:1987kr,allen2002,enquist1998}. %Previous studies have focused on t
This hypothesis is based on observations showing that the steady state
abundance, $N^*$, of a species is proportional to the inverse of individual
metabolism, such that
$N^*\propto M^{-3/4}/B_{0}$~\citep{allen2002,enquist1998}.  This relationship
implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been
shown to hold in both mammalian and vascular plant communities
\citep{Damuth:1987kr,allen2002,enquist1998}.  Figure \ref{fig:mass}a shows
that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of
organism sizes and Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly
constant over this same range.  This result is remarkable because it
illustrates that the steady state values of the NSM combined with the derived
timescales naturally give rise to energy equivalence.

@@

% \paragraph*{{\bf Maximum Body Size}}
Our model shows that energetic equivalence breaks down at large $M$ suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.  
%, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
With respect to predicted steady state densities, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occurs at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}). This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
In contrast, the $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$. 
This point predicts a more realistic upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is within an order of magnitude of the maximum body size observed in the North American mammalian fossil record~\citep{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\citep{Clauset:2009fh}.
It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\citep{Kempes:2012hy,Kempes:2016}. % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here

@@

The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca. $1.5\times10^7$g and the late Miocene \emph{Deinotherium} at ca. $1.74\times10^7$g ~\citep{Smith:2010p3442}.
Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \citep{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals explored by Alroy~\citep{Alroy:1998p1594}.
%as well as  \citep{Alroy:1998p1594,Clauset:2009fh}.
%This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals.\\

\noindent {\bf Discussion} \\
{\color{red} Complementary previous work has analyzed the tradeoffs associated with carnivory in terms of the energy expenditure on hunting and prey size, where a maximum carnivore size between 700 and 1,000 is predicted from considerations of energy intake going to zero \cite{Carbone:1999ju}. Similarly, it has been shown in the past that the largest body size within a particular lineage should be a function of the metabolic normalization constant and a critical death rate as supported by data. This approach, which consider the energetics associated with active metabolism, provides a complementary prediction for maximum size from energetics \cite{Okie:2013ju}, and also suggests that future connections between the scaling of fat and muscle mass should be systematically connected with $B_{0}$.
}

@@

Several efforts have shown how a partitioning of $B$ between growth and
maintenance purposes can be used to derive a general equation for both the
growth trajectories and growth rates of organisms ranging from bacteria to
metazoans
\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition 
$B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,$
% \begin{eqnarray}
% \label{balance}
% B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
% \end{eqnarray}
\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
the metabolic rate to support an existing unit of mass, and $m$ is the mass
of the organism at any point in its development.  This balance has the
general solution \citep{bettencourt,Kempes:2012hy}
\begin{eqnarray}
\label{m1}
\left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}},
\end{eqnarray}
where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass,
$a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically
(see the SI).  We now use this solution to define the timescale for
reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see
\citep{moses2008rmo} for a detailed presentation of these timescales). The
time that it takes to reach a particular mass $\epsilon M$ is given by the
timescale
\begin{equation}
\label{t1}
\tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)},
\end{equation}
where we will define values of $\epsilon$ to describe a set of rates within
our model.  The time to reproduce is given by
$t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
is the fraction of the asymptotic mass where an organism is reproductively
mature and should be close to one (typically
$\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}). The growth rate then
is $\lambda=\ln\left(\nu\right)/t_{\lambda}$ where $\nu$ is the number of
offspring produced, and for any constant value of $\epsilon_{\lambda}$, this
rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$
\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.

>>>>

Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
Many mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development until nutritional reserves can be accumulated~\citep{Mead:1989dt,Sandell:1990kw}.
Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\citep{Bulik:1999eo,Trites:2003cc}.
In the extreme case of unicellular organisms, nutrition directly controls growth to a reproductive state \citep{Glazier:2009hq,Kempes:2012hy}. The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the universality of the fundamental tradeoff between somatic and reproductive investment.

@@

% \noindent {\bf Energetic barriers to body size}
\noindent \paragraph*{{\bf Predicting Damuth's Law and body size limits.}} The NSM correctly predicts that species with smaller masses have larger
steady-state population densities (Fig.~\ref{fig:mass}a).  Similar
predictions have been made for carnivore populations using alternative
consumer-resource models \citep{DeLong:2012kw}.  Moreover, we show that the
NSM provides independent theoretical support for Damuth's Law and the energy equivalence
hypothesis\citep{Damuth:1987kr,allen2002,enquist1998,pedersen2017shallow}. {\color{red} Damuth's law shows that the  abundances of species, $N^{*}$, follows $N^*\propto M^{-0.78}$. The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}. As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}, however the actually equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations. Figure \ref{fig:mass}a shows
that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of
organism sizes and that $F^{*}+H^{*}$ closely matches the best fit to Damuth's data. Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly constant over this same range.  This result is remarkable because it
illustrates that the steady state values of the NSM combined with the derived
timescales naturally give rise to Damuth's law. While the previous metabolic studies supporting Damuth's law provided arguments for the value of the exponent \citep{allen2002}, these studies are only able to infer the intercept from the data. Our model predicts not only the exponent but also this intercept by explicitly including the resource dynamics and the parameters adjusting growth and consumption.

It should be noted the density relationships of individual clades follow more shallow scaling relationships than Damuth's law \cite{pedersen2017shallow}. In the context of our model, this suggests that future work may be able to anticipate these shifts from differences in the physiological parameters associated with each clade.}

@@

% \paragraph*{{\bf Maximum Body Size}}
Our model shows that energetic equivalence breaks down at large $M$ suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.  
%, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
With respect to predicted steady state densities, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occurs at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}). This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass of a male African elephant.
In contrast, the $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$. 
This point predicts a more realistic upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$ (g).
Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is within an order of magnitude of the maximum body size observed in the North American mammalian fossil record~\citep{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\citep{Clauset:2009fh}.
It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\citep{Kempes:2012hy,Kempes:2016}. {\color{red} It has also been shown that models of the energetic allometry of hunting and resting combined with foraging time predicts a maximum carnivore size between $7\times10^{5}$ and $1.1\times10^{6}$ (g) \cite{Carbone:1999ju,carbone2007costs}. {\color{red} Similarly, other past work has shown that the maximum body size within a particular lineage should scale with the metabolic normalization constant and depend on a critical death found to be constant from data \cite{Okie:2013ju}. This complementary approach is based on the balance between death and growth, and in connection with our model here suggests that future connections between the scaling of fat and muscle mass should be systematically connected with $B_{0}$ when comparing lineages.}} % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here

@@

The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca. $1.5\times10^7$g and the late Miocene \emph{Deinotherium} at ca. $1.74\times10^7$g ~\citep{Smith:2010p3442}.
Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \citep{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals explored by Alroy~\citep{Alroy:1998p1594}. {\color{red} It should be noted that the model of \cite{Brown:1993p708} predicts an optimal mammal size at intermediate mammalian sizes using reproductive maximization. This model considers the transition between hungry and full individuals as well, but our coupling to resources and an explicit treatment of storage shows that the largest mammals have an advantage in terms resource competition. Another subtlety is that the distributions in \cite{Brown:1993p708} show that intermediate mammal sizes have the most number of species, and it is important to recognize that our efforts consider total biomass where recent work shows that many ecosystems are dominated by the biomass of the large \cite{}.}
%as well as  \citep{Alroy:1998p1594,Clauset:2009fh}.
%This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.

While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals.\\

\noindent {\bf Discussion} \\

@@

Several efforts have shown how a partitioning of $B$ between growth and
maintenance purposes can be used to derive a general equation for both the
growth trajectories and growth rates of organisms ranging from bacteria to
metazoans
\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition 
$B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,$
% \begin{eqnarray}
% \label{balance}
% B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
% \end{eqnarray}
\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
the metabolic rate to support an existing unit of mass, and $m$ is the mass
of the organism at any point in its development.  This balance has the
general solution \citep{bettencourt,Kempes:2012hy}
\begin{eqnarray}
\label{m1}
\left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}},
\end{eqnarray}
where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass,
$a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically
(see the SI).  We now use this solution to define the timescale for
reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see
\citep{moses2008rmo} for a detailed presentation of these timescales). The
time that it takes to reach a particular mass $\epsilon M$ is given by the
timescale
\begin{equation}
\label{t1}
\tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)},
\end{equation}
{\color{red} where we will define values of $\epsilon$ to describe a variety of timescales, and related rates, within
our model.  For example, the rate of reproduction is given by the timescale to go from the birth mass to the adult mass. The time to reproduce is given by Equation \ref{t1} as $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
is the fraction of the asymptotic mass where an organism is reproductively
mature and should be close to one (typically
$\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}). Our reproductive rate, $\lambda$, is a specific rate, or the number of offspring produced per time per individual, defined as $\dot{F} = \lambda F$. In isolation this functional form gives population growth following $F\left(t\right) = F_{0}e^{\lambda t}$ which can be related to the reproductive timescale by assuming that when $t=t_{\lambda}$ it is also the case that $F=\nu F_{0}$, where $\nu-1$ is the number of offspring produced per reproductive cycle. Following this relationship the growth rate is given by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard relationship \cite{} and will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of $\epsilon_{\lambda}$
\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.}
